---
title: "ESM 206 Lab 5"
author: "Andrew Salvador"
date: "10/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives

- Getting counts fordifferent groups
- Use {lubridate} package to parse dates
- Find confidence intervals and do t-test with t.test() function
- Makeour first heatmap with geom_tile() function

```{r, include = FALSE}

# Attach packages

library(tidyverse)
library(here)
library(janitor)

```

```{r}
# Import lobster data 

lobster_abundance <- read_csv(here("data", "lobster_abundance.csv"), 
                              na = "-99999") %>% 
  clean_names()


```
Use the tidyr::uncount() function to expand a frequency variable, in this case lobster count

```{r}
lobster_tidy <- lobster_abundance %>% 
  uncount(lobster_count)
```

Now each lobster has its own row. This is tidy format. 

Now lets look at this data a bunch of ways. 

For the histogram, use facet_wrap ~site to make different graphs for each site. Use scales = "free" to make the scales to scale for each graph
```{r}

ggplot(lobster_tidy, aes(x = site, y = size_mm)) +
  geom_jitter(aes(color = site),
              width = 0.2,
              alpha = 0.3)

ggplot(lobster_tidy, aes(x = site, y = size_mm)) +
  geom_violin(aes(color = site),
              alpha = 0.3)


# Histogram

ggplot(lobster_tidy, aes(x = size_mm)) +
  geom_histogram(aes(fill = site)) +
  facet_wrap(~site, scales = "free")

# Now geom_qq

ggplot(lobster_tidy, aes(sample = size_mm)) +
  geom_qq(size = 0.5) +
  facet_wrap(~site, scales = "free")

```



