---
title: "ESM 206 Lab 5"
author: "Andrew Salvador"
date: "10/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives

- Getting counts fordifferent groups
- Use {lubridate} package to parse dates
- Find confidence intervals and do t-test with t.test() function
- Makeour first heatmap with geom_tile() function

```{r, include = FALSE}

# Attach packages

library(tidyverse)
library(here)
library(janitor)

```

```{r}
# Import lobster data 

lobster_abundance <- read_csv(here("data", "lobster_abundance.csv"), 
                              na = "-99999") %>% 
  clean_names()


```
Use the tidyr::uncount() function to expand a frequency variable, in this case lobster count

```{r}
lobster_tidy <- lobster_abundance %>% 
  uncount(lobster_count)
```

Now each lobster has its own row. This is tidy format. 

Now lets look at this data a bunch of ways. 

For the histogram, use facet_wrap ~site to make different graphs for each site. Use scales = "free" to make the scales to scale for each graph
```{r}

ggplot(lobster_tidy, aes(x = site, y = size_mm)) +
  geom_jitter(aes(color = site),
              width = 0.2,
              alpha = 0.3)

ggplot(lobster_tidy, aes(x = site, y = size_mm)) +
  geom_violin(aes(color = site),
              alpha = 0.3)


# Histogram

ggplot(lobster_tidy, aes(x = size_mm)) +
  geom_histogram(aes(fill = site)) +
  facet_wrap(~site, scales = "free")

# Now geom_qq

ggplot(lobster_tidy, aes(sample = size_mm)) +
  geom_qq(size = 0.5) +
  facet_wrap(~site, scales = "free")

```

Now we are going to use the 'lubridate' package to parse dates and times (we are going to make it date class)
Use the explicit version of this package because it may show up in other packages

First I'm going to add a new column with 'mutate()', that contains my date as an actual date
mdy is because the date is currently written as month/day/year. The lubridate function automatically puts the date into standard ISO format (YYYY-MM-DD)

```{r}
lobster_date <- lobster_tidy %>% 
  mutate(
    date_new = lubridate::mdy(date)
  )


```

What if I want to sort by month

Now we will Parse 'Date' column to get the different pieces (month, year) separated.
If object is in 'date' class, use lubridate::month function and it will pull the month. Use label = TRUE to change it to month name
```{r}
lobster_parse_date <- lobster_date %>% 
  mutate(
    obs_month = lubridate::month(date_new, label = TRUE),
    obs_year = lubridate::year(date_new)
  )
  
  
```


Now we are going to count lobsters by different groupings using 'count' function
Let's say I want to count the number of lobsters by year and month

The 'count' function does a 'group_by' then 'summarize', then find the length, then lastly it ungroups  the data

'count' is combination of 'group_by' and 'tally'

```{r}

lobster_ym <- lobster_parse_date %>% 
  count(obs_year, obs_month) 

lobster_ysite <- lobster_parse_date %>% 
  count(site, obs_year)

  
#Now...
# group by + summarize + n() to create a new table with mean size, and standard deviation function and sample size (n)

lobster_summary <- lobster_parse_date %>% 
  group_by(site) %>% 
  summarize(
    mean_size = mean(size_mm, na.rm = TRUE),
    sd_size = sd(size_mm, na.rm = TRUE),
    sample_n = n()
  )
  


```


Next we'll do some confidence intervels and t-tests

Use the 't.test()' function to find confidence intervals and perform t-tests

I wanna find confidence intervals for lobster sizes at isla vista
use 'pull()' function to pull out the size_mm column into just one vector.
Calling an indivudal column with using dataframe$columnname will do the same thing as vector
```{r}
ivee_lobsters <- lobster_tidy %>% 
  filter(site == "IVEE") %>% 
  pull(size_mm)

t.test(ivee_lobsters) # 95% confidence interval that in 95 percent in other samples I theoretically take IO would expdct teh sample mean to fall within this range. 

```

default in onesample t test
asking ifwe have enough evidence to think that actual value significnatly different than zero

Now....
Two sample t-test:
Is there a significant difference in mean lobster lengths between Naples Reef and Mohawk Reef?


```{r}

# method one (more tedious)

# pull the vectors

napl_sample <- lobster_tidy %>% 
  filter(site == "NAPL") %>% 
  pull(size_mm)

mohk_sample <- lobster_tidy %>% 
  filter(site == "MOHK") %>% 
  pull(size_mm)

# Now t test
mohk_napl_ttest <- t.test(napl_sample, mohk_sample)

mohk_napl_ttest

```
p value means that there is only a 2.2 x 10 to the minus 16th probability that i could have taken 2 samples from poulations with the same mean and found by random chance, means of my groups that are at least that different 

p value is thep robability. So in our test, the probability is super low so we are going to assume thatwe took samples from populations that actually havedifferent means

inline referencing: 

Mean lobster size differed significantly between Mohawk and Naples reef (t(`r mohk_napl_ttest$parameter`) = 19.849)

Maybe an easier way to do a 2 sample t.test
t.test function only works with 2 groups, so we need to make the dataframe 2 groups before doing the t test
```{r}
# METHOD 2

#create data frame that only has 2 groups I want to compare

lobster_2sample <- lobster_tidy %>% 
  filter(site %in% c("NAPL", "MOHK"))

ttest_2 <- t.test(size_mm ~ site, data = lobster_2sample)

ttest_2


```
Now lets make a geom_tile heatmap

```{r}

ggplot(lobster_ysite, aes(x = obs_year,
                          y = site)) +
  geom_tile(aes(fill = n))



```


